|

### 2020-06-30 paper.Rmd
More work on introduction.

In this work we address this piling in low-dimensional linear projections, which is in particular a barrier for viewing high-dimensional distributions with tour methods [@As85,@BCAH05]. Tours show interpolated sequences of low-dimensional projections of the data. When exploring a dataset with a tour we look at the distribution from different sides and want to discover features that may be visible from a specific viewing angle. However, these features may easily be hidden if most of the observations pile up near the center.  XXX talk about few outliers taking up large part of the plotting space

@@

# Projected volume

The seemingly contradicting behaviour between the low density of points near the center in the high-dimensional space, and the piling in low-dimensional projections can be understood by considering the projected volume across the low-dimensional projection. Assuming that the total distance from the center is constrained, we can think of the data being distributed in a hypersphere, which has the additional advantage of being rotation invariant. The radial dependence of the projected volume was described in @Laa:2020wkm via a ratio of the total volume of a $p$ dimensional sphere of radius $R$, $V(R, p)$ and its projected volume within a 2D radius $r$, $V_{2D}(r, R, p)$,
\begin{equation}
f_{2D} (r, p, R) = \frac{V_{2D}(r, R, p)}{V(R, p)} = 1 - \left(1-\left(\frac{r}{R}\right)^2\right)^{p/2}.
\end{equation}
This ratio is of particular interest because it gives the 2D radial cumulative distribution function (CDF) of points when assuming a uniform distribution within the $p$ dimensional hypersphere. We can compare the radial CDF $f_{2D} (r, p, R)$ to the radial CDF in the full hypersphere,
\begin{equation}
f_{pD} (r, p, R) = \frac{V(r, p)}{V(R, p)} = \left({\frac{r}{R}}\right)^p.
\end{equation}
We show this comparison for selected values in Fig. \ref{fig:cdf}, demonstrating the opposing behaviour as $p$ increases. While $f_{pD}$ concentrates more and more towards large values of $r/R$, $f_{2D}$ has the opposite behaviour, i.e. the volume projected to two dimensions is concentrated near the center of the disk, since we are integrating the large number of orthogonal directions.

@@

# Implementation

@@

# Applications

@@

# Discussion

>>>>

In this work we address the piling in low-dimensional linear projections, which is in particular a barrier for viewing high-dimensional distributions with tour methods [@As85,@BCAH05]. Tours show interpolated sequences of low-dimensional projections of the data. When exploring a dataset with a tour we look at the distribution from different sides and want to discover features that may be visible from a specific viewing angle. However, these features may easily be hidden if most of the observations pile up near the center, while a large fraction of the plotting canvas will only show sparse outlying points. Here we suggest to redistribute the projected points such that equal volume in an original hypersphere will be mapped onto equal areas on a 2D projection plane. This redistribution is achieved through a non-linear transformation of the projected radius of the observations, and essentially results in zooming in near the center of the distribution, while pushing points towards the outer radius of the projected circle if they are further out.

The paper is structured as follows: the radial transformation is described in Section \ref{sec:method}, the implementation as a new display function for the tourr package is described in Section \ref{sec:implementation}, and we show applications in Section \ref{sec:application}. We conclude in Section \ref{sec:concl}.

@@

# Radial transformation {#sec:method}

As discussed in the introduction, the aim of the radial transformation is to redistribute the projected points, such that equal volume in the original ($p$-dimensional) space is projected onto equal area in a 2D projection. Note that it is straightforward to generalise this for projections onto $d$-dimensional projections in the same way.

The radial dependence of the projected volume was described in @Laa:2020wkm via a ratio of the total volume of a $p$ dimensional sphere of radius $R$, $V(R, p)$ and its projected volume within a 2D radius $r$, $V_{2D}(r, R, p)$,
\begin{equation}
f_{2D} (r, p, R) = \frac{V_{2D}(r, R, p)}{V(R, p)} = 1 - \left(1-\left(\frac{r}{R}\right)^2\right)^{p/2}.
\end{equation}
This ratio is of particular interest because it gives the 2D radial cumulative distribution function (CDF) of points when assuming a uniform distribution within the $p$ dimensional hypersphere. We can compare the radial CDF $f_{2D} (r, p, R)$ to the radial CDF in the full hypersphere,
\begin{equation}
f_{pD} (r, p, R) = \frac{V(r, p)}{V(R, p)} = \left({\frac{r}{R}}\right)^p.
\end{equation}
We show this comparison for selected values in Fig. \ref{fig:cdf}, demonstrating the opposing behaviour as $p$ increases. While $f_{pD}$ concentrates more and more towards large values of $r/R$, $f_{2D}$ has the opposite behaviour, i.e. the volume projected to two dimensions is concentrated near the center of the disk, since we are integrating the large number of orthogonal directions.

@@

# Implementation {#sec:implementation}

@@

# Applications {#sec:application}

@@

# Discussion {#sec:concl}
